Water discharge quantity has a variable impact on sediment transport capacity in subglacial channels

### 2023-09-30 Delaneyetal2023b.tex
Now added w*tau to table too
Scaling-text slowly getting there

\newpage
\section{Mauro's bit about shear stress}
\label{sect:scaling}
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options. This is common when trying to add a PDF figure without pdflatex.
% \begin{figure}
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
%\caption{caption}
%\label{pdffiguresample}
%\end{figure}
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendices go here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}
\end{document}

>>>>

\newpage
\section{Scalings for sediment transport in different drainage regimes}
\label{sect:scaling}

\mauro{Notation: either use $H$ and $W$ for the river, or otherwise, somehow free $h$. Below I use $H$ and $W$ for now.}

Here we look at three kind of drainage regimes, namely, rivers, steady-state R-channels, and pipe-flow (i.e. R-channels which do not have time to adjust their size to discharge conditions).  We calculate how in these regimes shear stress and sediment flux scale with respect to a given discharge for three different transport formulas: MPM \cite{meyer1948}, EH \cite{engelund1967}, and Bagnold \cite{Bagnold1980}.

A few basic hydraulic relations are used in the following.  The Darcy-Weisbach equation can be stated
\begin{equation}
  \label{eq:DW}
  \Psi \propto f\frac{Q^2}{D_h S^2},
\end{equation}
with $\Psi$ the head gradient and friction factor $f$.  The shear stress and stream power are, respectively,
\begin{equation}
    \label{eq:tau-omega}
  \tau \propto f v^2 = f \left(\frac{Q}{S}\right)^2, \quad  \Omega \propto \Psi Q.
\end{equation}
%
Sediment discharge is given by the MPM, EH and Bagnold formulations as, respectively,
\begin{equation}
  Q_s \propto w\, \tau^{3/2}, \quad Q_s \propto w\, \tau^{5/2}, \quad Q_s \propto w\, \Omega^{3/2}
\end{equation}
for conditions well above transport onset threshold.

We assume a wide river with depth $H$, such that $W \gg H$ and $D_h\approx 4H$, that has a constant head gradient ($\Psi$) given by the topography and we assume that its width can be approximated by a relation $w \propto Q^\alpha$ with $\alpha\in [0,1]$; typically $\alpha \approx 1/2$, $\alpha=0$ or $\alpha=1$ correspond a river of constant width or depth, respectively.

We assume that a steady state R-channel has a constant $\Psi$ (approximated by the gradient of the Shreve \citeyear{shreve1972} potential) and that $S$ adjusts such that it is in steady state.  Conversely, pipe flow has $S$ fixed and $\Psi$ adjusts.
We assume that for pipe flow, which occurs when a R-channel is subjected to rapid discharge variations such that it cannot adjust its size significantly, the cross-sectional area $S$ is fixed and that $\Psi$ adjusts.  For both R-channel and pipe flow, we assume that $D_h \propto S^{1/2}$.

With these assumptions the Darcy-Weisbach equation~\eqref{eq:DW} can be solved for the not-fixed quantity: $H$ for river, $S$ for R-channel, and $\Psi$ for pipe-flow.  Then, using equations~\eqref{eq:tau-omega}, the shear stress and stream power can be calculated.  These resulted are summarised in Table~\ref{tab:eqs1}.
Now sediment transport can be calculated for all regimes and all transport formulas.  The results of the nine combinations are presented in Table~\ref{tab:Qs} and also in Table~\ref{tab:Qs2} where the fractions in the exponents are approximately given by decimal numbers for ease of comparison.

\begin{table}
  \caption{}
  \label{tab:eqs1}
  \small
\begin{tabular}{llllll}
Regime & Fixed & Determined via D-W & Additional relations & Shear stress & Stream p.\\[0pt]
 &  &  &  & \(\tau \propto\) & \(\Omega \propto\)\\[0pt]
\hline
River & \(\Psi\) & \(H \propto f^{1/3}\, Q^{2/3-2\alpha/3} \, \Psi^{-1/3}\) & \(W\,\propto Q^\alpha\), \(S=WH\), \(D_h=H\) & \(f^{1/3} Q^{2/3-2\alpha/3}  \Psi^{2/3}\) & \(Q \Psi\)\\[0pt]
R-ch. & \(\Psi\) & \(S\, \propto f^{2/5}\, Q^{4/5} \, \Psi^{-2/5}\) & \(D_h\propto S^{1/2}\) & \(f^{1/5} Q^{2/5} \, \Psi^{4/5}\) & \(Q \Psi\)\\[0pt]
Pipe & \(S\) & \(\Psi \propto f \, Q^2\, S^{-5/2}\) & \(D_h\propto S^{1/2}\) & \(f Q^2 S^{-2}\) & \(f Q^3 S^{-5/2}\)\\[0pt]
\end{tabular}
\end{table}

\begin{table}
  \caption{
    \mauro{That negative exponent of $f$ for Bagnold-River is pretty crazy! Not sure what that is about.}
    }
  \label{tab:Qs}
\begin{tabular}{lllll}
 & Width*\(\tau\) & MPM & EH & Bagnold\\[0pt]
 & \(W\, \tau\) & \(Q_s \propto W\, \tau^{3/2}\) & \(Q_s \propto W\, \tau^{5/2}\) & \(Q_s \propto W\, \Omega^{3/2} H^{-2/3}\)\\[0pt]
\hline
River & \(f^{1/3}\, Q^{2/3+\alpha/3}\,  \Psi^{2/3}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{5/6}\, Q^{5/3 - 2\alpha/3} \, \Psi^{5/3}\) & \(f^{-2/9}\, Q^{13\alpha/9 - 19/18} \, \Psi^{31/18}\)\\[0pt]
R-ch. & \(f^{2/5}\, Q^{4/5} \, \Psi^{3/5}\) & \(f^{1/2}\, Q \, \Psi\) & \(f^{7/10}\, Q^{7/5}\, \Psi^{9/5}\) & \(f^{1/15}\, Q^{49/30}\, \Psi^{43/30}\)\\[0pt]
Pipe & \(f \, Q^2 \, S^{-1}\) & \(f^{3/2}\, Q^3 \, S^{-5/2}\) & \(f^{5/2}\, Q^5\, S^{-9/2}\) & \(f^{3/2} \, Q^{9/2} \, S^{-43/12}\)\\[0pt]
\end{tabular}
\end{table}

\mauro{TODO: make third table C3.  Discuss results briefly.}

\end{document}
